---
title: "Week 7 Problems"
author: "Cassandra"
date: "5/26/2020"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, autodep = TRUE)
```

```{r, message = FALSE}
library(tidyverse)
library(nycflights13)
library(cowplot) 
theme_set(theme_cowplot())
library(hexbin)
```

## Chapter 13 

### 13.2.1 

**1. Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?**
```{r}
head(airports)
head(flights)
```

You would need the location of each airport, so `lat` and `lon` from `airports` along with `faa`. You would also need the `origin` and `dest` from `flights`, which corresponds with `faa` from `airports`. You would need to combine `flights` and `airports`. 

**2. I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?**     
```{r}
head(weather)
head(airports)
```

`airports` connects to `weather` through `faa`, which corresponds to `origin` in `weather`. This should be portrayed as an arrow between `faa` and `origin`. 

**3. `weather` only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with `flights`?**      

`dest` would also be defined with `flights`. 

**4. We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?** 

You could make a `holidays` data frame that includes `name`, `month`, and `day`. The primary keys would be `month` and `day`, which is how it would connect to `flights` and `weather`. 

### 13.3.1

**1. Add a surrogate key to `flights`.**    

```{r}
flights2 <- flights %>% mutate(
  key = row_number()
)

flights2 %>% select(key, everything()) %>% head()
```

**2. Identify the keys in the following datasets:**    

  1. Lahman::Batting,
  2. babynames::babynames
  
**(You might need to install some packages and read some documentation.)** 

```{r}
library(Lahman)
head(Batting)
Batting %>% count(playerID, yearID, stint) %>% filter(n > 1)
```
```{r}
library(babynames)
head(babynames)
babynames %>% count(year, sex, name) %>% filter(n > 1)
```

### 13.4.6 

**1. Compute the average delay by destination, then join on the `airports` data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:**  
```
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```
**(Don’t worry if you don’t understand what semi_join() does — you’ll learn about it next.)**    
**You might want to use the size or colour of the points to display the average delay for each airport.**     
```{r}
# compute average delay by destination
avg_delay <- flights %>% 
  group_by(dest) %>% 
  summarise(
    arrival = mean(arr_delay, na.rm = TRUE),
    departure = mean(dep_delay, na.rm = TRUE))

# join on airports 
avg_delay %>% 
  left_join(airports, c("dest" = "faa")) %>% 
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point(aes(color = arrival)) +
    coord_quickmap()

avg_delay %>% 
  left_join(airports, c("dest" = "faa")) %>% 
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point(aes(color = departure)) +
    coord_quickmap()
```

**2. Add the location of the origin *and* destination (i.e. the `lat` and `lon`) to `flights`.**  

```{r}
location <- airports %>% 
  select(faa, lat, lon)

flights %>% 
  left_join(location, c("dest" = "faa")) %>% 
  rename(dest_lat = lat,
         dest_lon = lon) %>% 
  left_join(location, c("origin" = "faa")) %>% 
  rename(origin_lat = lat,
         origin_lon = lon) %>% 
  select(origin, origin_lat, origin_lon, dest, dest_lat, dest_lon, everything())
```

### 13.5.1 

**2. Filter flights to only show flights with planes that have flown at least 100 flights.**    

```{r}
most_flown <- flights %>% 
  filter(tailnum != "NA") %>% 
  count(tailnum, sort = TRUE) %>% 
  filter(n >= 100)

flights %>% 
  semi_join(most_flown)
```

**4. Find the 48 hours (over the course of the whole year) that have the worst delays. Cross-reference it with the `weather` data. Can you see any patterns?**      

```{r}
worst_delay <- flights %>% 
  group_by(year, month, day, hour) %>% 
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>% 
  arrange(desc(delay)) %>% 
  head(48)

weather_delay <- worst_delay %>% 
  left_join(weather) %>% 
  select(year, month, day, hour, delay, origin, temp, wind_speed, precip, visib, everything())

summary(weather_delay)
summary(weather)
```

## Chapter 28

### 28.2.1 

**1. Create one plot on the fuel economy data with customised `title`, `subtitle`, `caption`, `x`, `y`, and `colour` labels.**      

```{r}
ggplot(mpg) + 
  geom_point(aes(cty, hwy, color = class)) + 
  labs(
    title = "City mpg and highway mpg are highly correlated",
    subtitle = "SUV and pickups have the worst mpg",
    caption = "Data from fueleconomy.gov",
    x = "City Fuel Economy (mpg)",
    y = "Highway Fuel Economy (mpg)",
    color = "Car Class"
  )
```

### 28.4.4

**1. Why doesn’t the following code override the default scale?**    
```
ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_colour_gradient(low = "white", high = "red") +
  coord_fixed()
``` 
I'm guessing the gradient scaling is actually controlled by `fill` and not `color`? 

```{r}
df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)

ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_fill_gradient(low = "white", high = "red") +
  coord_fixed()
```

**4. Use `override.aes` to make the legend on the following plot easier to see.**    

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_point(aes(colour = cut), alpha = 1/20) + 
  guides(color = guide_legend(override.aes = list(alpha = 1)))
```

