---
title: "Week 6 Problems"
author: "Cassandra"
date: "5/25/2020"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, autodep = TRUE)
```

```{r, message = FALSE}
library(tidyverse)
```

## Chapter 12

### Exercises 

#### 12.2.1 

**1. Using prose, describe how the variables and observations are organised in each of the sample tables.**   
In `table1` each variable has its own column. In `table2` some variables share columns. In `table3` some values share cells. In `table4a` and `table4b` some variables are spread across multiple columns and some are spread across multiple tables. 

**2. Compute the `rate` for `table2`, and `table4a` + `table4b`. You will need to perform four operations:**    

  1. Extract the number of TB cases per country per year.
  2. Extract the matching population per country per year.
  3. Divide cases by population, and multiply by 10000.
  4. Store back in the appropriate place.
  
**Which representation is easiest to work with? Which is hardest? Why?**    

```{r}
# for table 2
## extract cases
cases <- table2 %>% filter(type == "cases")
## extract population
pop <- table2 %>% filter(type == "population")
## calculate rate 
rate <- tibble(
  country = rep(unique(table2$country), times = rep_len(length(unique(table2$year)), length.out = length(unique(table2$country)))),
  year = rep(unique(table2$year), times = length(unique(table2$country))),
  type = "rate",
  count = cases$count / pop$count * 10000
)

## store in table2
table2_mod <- rbind(table2, rate) %>% 
  arrange(country, year, type) 
table2_mod 
```

```{r}
# for table 4a + 4b
## cases and population are already 'extracted'
## calculate rate 
table4c <- tibble(
  country = table4a$country,
  `1999` = table4a$`1999` / table4b$`1999` * 10000,
  `2000` = table4a$`2000` / table4b$`2000` * 10000
)
table4c
```

I guess here `table4` is easier to work with, but it frustratingly does not have all the data conveniently stored together. Conversely, `table2` does have everything stored together but is a pain to work with. Tidy data fixes these problems! 

**3. Recreate the plot showing change in cases over time using `table2` instead of `table1`. What do you need to do first?**      

```{r}
# need to filter
table2 %>% filter(type == "cases") %>% 
  ggplot(aes(year, count)) + 
  geom_line(aes(group = country), color = "grey50") + 
  geom_point(aes(color = country)) + 
  ylab("cases")
```

#### 12.3.3 

**1. Why are `pivot_longer()` and `pivot_wider()` not perfectly symmetrical?**   
**Carefully consider the following example:**     
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return")
``` 
**(Hint: look at the variable types and think about column names.)**

Looks like `year` has lost its variables type. So I suppose these are not perfectly symmetrical because column names don't have a type associated with them, so as soon as variables are moved into columns their type is lost. 

**`pivot_longer()` has a `names_ptype argument`, e.g. `names_ptype = list(year = double())`. What does it do?**    

```{r}
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return", names_ptype = list(year = double()))
```

This argument specifies what type your 'new' variable should be. Handy!

**2. Why does this code fail?**   
```
table4a %>% 
  pivot_longer(c(1999, 2000), names_to = "year", values_to = "cases")
#> Error: Can't subset columns that don't exist.
#> ✖ The locations 1999 and 2000 don't exist.
#> ℹ There are only 3 columns.
```

This probably fails because "1999" and "2000" are non-syntactic names but are not surrounded by `` ` `` in the previous code. 
```{r}
table4a %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
```

**3. What would happen if you widen this table? Why? How could you add a new column to uniquely identify each value?** 
```{r}
people <- tribble(
  ~name,             ~names,  ~values,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
```

```{r}
people %>% pivot_wider(names_from = names, values_from = values)
```

R actually lets you know that there are multiple values for one observation (because there are two different people named "Phillip Woods"). You could add a unique ID number for each person. 
```{r}
people %>% 
  mutate(ID = c(1, 1, 2, 3, 3)) %>% 
  pivot_wider(names_from = names, values_from = values)
```

**4. Tidy the simple tibble below. Do you need to make it wider or longer? What are the variables?**   
```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
```

Need to make this longer because values of a variable are column names. The variables are `pregnant`, `sex`, and `count`. 
```{r}
preg %>% pivot_longer(c("male", "female"), names_to = "sex", values_to = "count")
```

#### 12.4.3

**1. What do the `extra` and `fill` arguments do in `separate()`? Experiment with the various options for the following two toy datasets.**    
```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
```

`extra` determines what happens with too many observations and `fill` determines what happens with too few observations. 
```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "merge")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "left")
```

**2. Both `unite()` and `separate()` have a `remove` argument. What does it do? Why would you set it to `FALSE`?** 
The `remove` argument will remove the original input column by default. You would set this to `FALSE` if you want to keep the original input in addition to the newly generated output(s).   
```{r}
table3 %>% separate(year, into = "century", sep = 2, remove = FALSE)
```

**3. Compare and contrast `separate()` and `extract()`. Why are there three variations of separation (by position, by separator, and with groups), but only one unite?**   

Seems like `separate()` generally splits on a separator or position while `extract()` uses regular expressions to separate by groups. There are a lot of different ways that I can imagine wanting to separate values, but uniting values seems to have a more universal mechanism/output. 

#### 12.6.1

```{r}
who_tidy <- who %>%
  pivot_longer(
    cols = new_sp_m014:newrel_f65, 
    names_to = "key", 
    values_to = "cases", 
    values_drop_na = TRUE
  ) %>% 
  mutate(
    key = stringr::str_replace(key, "newrel", "new_rel")
  ) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```

**3. I claimed that `iso2` and `iso3` were redundant with `country`. Confirm this claim.**    
```{r}
who %>% select(1:3) %>% 
  unique() %>% 
  count(country) %>% 
  filter(n > 1)
```

**4. For each country, year, and sex compute the total number of cases of TB. Make an informative visualisation of the data.**    
```{r}
who_tidy %>% 
  count(country, year, sex, wt = cases) %>% 
  ggplot(aes(year, n, color = country)) + 
  geom_point() + 
  geom_line() + 
  facet_wrap(~ sex) + 
  theme(legend.position = "none")
```

There are so many countries that this isn't great, but it does look like males may have higher frequency of TB? 

## Joins Tutorial

### Types 

* `inner_join`: only keeps rows with a key shared in both datasets    
* `left_join`: keeps all rows from left dataset and only keeps rows from right dataset that have a matching key with the left dataset    
* `right_join`: keeps all rows from right dataset and only keeps rows from left dataset that have a matching key with the right dataset    
* `full_join`: keeps all rows from both datasets   

### Keys    

If the key is under a different column name in the two datasets, you can specify that the two different names are equivalent with `by`.   