---
title: "Week 3 Problems"
author: "Cassandra"
date: "5/1/2020"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, autodep = TRUE)
```

```{r, message = FALSE}
library(tidyverse)
library(nycflights13)
```

## Chapter 4

### Exercises

#### 4.4

**1. Why does this code not work?**    
```
my_variable <- 10
my_varıable
```

The second line has another character in place of the "i".   

**2. Tweak each of the following R commands so that they run correctly:**   
```
ggplot(dota = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

fliter(mpg, cyl = 8)
filter(diamond, carat > 3)
```
```{r}
# dota to data 
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# fliter to filter, = to ==
filter(mpg, cyl == 8)

# diamond to diamonds
filter(diamonds, carat > 3)
```

**3. Press `Alt` + `Shift` + `K`. What happens? How can you get to the same place using the menus?**    
This brings up a cool list of keyboard shortcuts for R. You can access this using the menus by clicking `Help` and `Keyboard Shortcuts Help`.   

## Chapter 5 

### Exercises 

#### 5.2.4

**1. Find all flights that**   
  
**1.1 Had an arrival delay of two or more hours**    
```{r}
filter(flights, arr_delay >= 120) %>% 
  ggplot() + 
  geom_histogram(aes(arr_delay)) + 
  xlim(0, 500)
```

**1.4 Departed in summer (July, August, and September)**    
```{r}
filter(flights, month %in% c(7:9)) %>% 
  ggplot() + 
  geom_histogram(aes(month))
```

**1.5 Arrived more than two hours late, but didn't leave late**   
```{r}
filter(flights, arr_delay > 120, dep_delay <= 0) %>% 
  ggplot() + 
  geom_point(aes(arr_delay, dep_delay))
```

**1.7 Departed between midnight and 6 am (inclusive)**    
```{r}
filter(flights, dep_time <= 600 | dep_time == 2400) %>% 
  ggplot() + 
  geom_histogram(aes(dep_time))
```

**2. Another useful `dplyr` filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?**    
`between()` pulls out values equal to or between two boundary values. This could certainly be used for problem 1.4 above.  
```{r}
filter(flights, between(month, 7, 9)) %>% 
  ggplot() + 
  geom_histogram(aes(month))
```

**3. How many flights have a missing `dep_time`? What other variables are missing? What might these rows represent?**    
```{r}
is.na(flights$dep_time) %>% sum()
summary(flights)
```

Looks like `dep_delay`, `arr_time`, `arr_delay`, and `air_time` all have missing values. These rows could be cancelled flights?    

#### 5.3.1

**1. How could you use `arrange()` to sort all missing values to the start? (Hint: use `is.na()`).**   
```{r}
# if NA, then = 1 --> desc() so NA (1) is on top 
arrange(flights, desc(is.na(dep_time))) %>% head()
```

**2. Sort `flights` to find the most delayed flights. Find the flights that left earliest.**   
```{r}
# most delayed
arrange(flights, desc(arr_delay)) %>% print(n = 6, width = Inf)

# earliest departure
arrange(flights, dep_delay) %>% head()
```

**3. Sort `flights` to find the fastest (highest speed) flights.**    
```{r}
# can sort to find fastest flights in terms of shortest 
arrange(flights, air_time) %>% print(n = 6, width = Inf)

# doesn't seem like any of the existing variables provide information on speed though...
mutate(flights,
       speed = distance / air_time) %>% 
  arrange(desc(speed)) %>% 
  print(n = 6, width = Inf)
```

**4. Which flights travelled the farthest? Which travelled the shortest?**     
```{r}
# farthest distance 
arrange(flights, desc(distance)) %>% print(n = 6, width = Inf)

# shortest distance
arrange(flights, distance) %>% print(n = 6, width = Inf)
```

#### 5.4.1 

**1. Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from `flights`.**     
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay) %>% head()
select(flights, dep_time:arr_delay, -contains("sched")) %>% head()
select(flights, contains(c("dep", "arr_")), -contains("sched")) %>% head()
select(flights, starts_with(c("dep", "arr"))) %>% head()
select(flights, ends_with(c("time", "delay")), -starts_with(c("sched", "air"))) %>% head()
select(flights, 4:9, -contains("sched")) %>% head()
select(flights, -(1:3), -(10:19), -5, -8) %>% head()
```

**2. What happens if you include the name of a variable multiple times in a `select()` call?**    
```{r}
select(flights, air_time, air_time, distance) %>% head()
```

**3. What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?** 
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars)) %>% head()
```

Looks like `one_of()` just selects for the variable names included within the vector. This would be useful if you have many lines of code or graphs working with a specific subset of columns and you wanted to easily change the column subset for all those lines by changing only the initial variable vector. Also, `one_of()` is now replaced by `any_of()` and `all_of()`. 

**mutat4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?**    
```{r}
select(flights, contains("TIME"))
```

This result actually does surprise me because R is usually very case-sensitive. Looks like `select` is set to ignore case by default for matching. You can change the default with `ignore.case = FALSE`.    
```{r}
select(flights, contains("TIME", ignore.case = FALSE))
```

#### 5.5.2   

**1. Currently `dep_time` and `sched_dep_time` are convenient to look at, but hard to compute with because they're not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.**     
```{r}
mutate(flights, 
       dep_min = (dep_time %/% 100) * 60 + dep_time %% 100,
       sched_dep_min = (sched_dep_time %/% 100) * 60 + sched_dep_time %% 100
       ) %>% 
  print(n = 6, width = Inf)
```

**2. Compare `air_time` with `arr_time - dep_time`. What do you expect to see? What do you see? What do you need to do to fix it?**   
```{r}
select(flights, air_time, arr_time, dep_time) %>% 
  mutate(
    flight_time = arr_time - dep_time
  ) %>% 
  head()
```

I guess we expect these to be the same? Instead, these are different values because `arr_time` and `dep_time` are in a strange hour-minute format and not a format useful for calculations. This could be fixed by adjusting `arr_time` and `dep_time` into another format, like hours since midnight.    

**3. Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you expect those three numbers to be related?**  
```{r}
select(flights, dep_time, sched_dep_time, dep_delay) %>% 
  mutate(
    dep = sched_dep_time + dep_delay
  ) %>% 
  head()
```

Generally the `dep_time` should be the `sched_dep_time` + `dep_delay`, but this doesn't work when the `dep_time` and `sched_dep_time` are across different hours. 

**4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().**    
```{r}
mutate(flights, 
       delay_rank = min_rank(desc(arr_delay))) %>% 
  arrange(delay_rank) %>% 
  print(n = 10, width = Inf)
```

**5. What does `1:3 + 1:10` return? Why?**   
```{r}
1:3 + 1:10
```

A vector of numbers where 1:3 is added to 1:10 in a way that 1:3 is recycled, so 1 is added to 1, 4, 7, and 10. This occurs because R recycles vectors and 1:3 is shorter than 1:10. 

#### 5.6.7

**5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about `flights %>% group_by(carrier, dest) %>% summarise(n())`)**
```{r}
#flights %>% group_by(carrier, dest) %>% summarise(n()) %>% View()
```

Hmm.. well some airports are only served by one carrier and some carriers only serve one airport. Some of these carriers/airports also have a low number of flights and would be more susceptible to variation. 

```{r}
# remove cancelled flights 
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

# simple determination of worst delays
not_cancelled %>% 
  group_by(carrier) %>% 
  summarise(mean(arr_delay)) %>% 
  arrange(desc(`mean(arr_delay)`))
```

Okay let's look at some of these carriers in more detail. We're supposed to be looking at worst delays, but the negative delay average of HA and AS seems interesting. 
```{r}
not_cancelled %>% filter(carrier %in% c("HA", "AS")) %>% 
  ggplot(aes(distance, arr_delay)) + 
  geom_point(aes(color = dest)) + 
  facet_wrap(~ carrier)
```

Ah, so each of these carriers only flies to one destination. Might make it easier to have predictable arrival times. 

What about the worst delay carriers in comparison? 
```{r}
not_cancelled %>% filter(carrier %in% c("F9", "FL")) %>% 
  ggplot(aes(distance, arr_delay)) + 
  geom_point(aes(color = dest)) + 
  facet_wrap(~ carrier)
```

Looks like the carrier with the worst average delays, F9, also only flies to one destination. Do other carriers fly to Denver?   
```{r}
not_cancelled %>% filter(dest == "DEN") %>% 
  ggplot(aes(carrier, arr_delay)) + 
  geom_point(aes(color = carrier))
```

Looking at the other carriers that fly to Denver, it looks like F9 doesn't have particularly worse delays than anyone else. F9 just has one outlier flight that is dragging the mean delay up. 
