---
title: "Week 5 Problems"
author: "Cassandra"
date: "5/18/2020"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, autodep = TRUE)
```

```{r, message = FALSE}
library(tidyverse)
```

## Chapter 10 

You can use `.` in a pipe as a placeholder!   
```{r}
diamonds %>% filter(carat > 4) %>% .$carat
```

### Exercises 

#### 10.5 

**1. How can you tell if an object is a tibble? (Hint: try printing `mtcars`, which is a regular data frame).** 
```{r}
print(mtcars)
print(diamonds)
```

**2. Compare and contrast the following operations on a `data.frame` and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?**   
```{r}
# make test data frame
df <- data.frame(abc = 1, xyz = "a")
# make test tibble
tib <- tibble(abc = 1, xyz = "a")
```

```{r}
df$x
tib$x
```
The data frame lets you do this even though it doesn't match the column name, which may lead to you working with the wrong column on accident. 

```{r}
df[, "xyz"]
tib[, "xyz"]
```
With a tibble the output is also a tibble and is visually more pleasing. I kind of like how the data frame gives you the factor levels though. 

```{r}
df[, c("abc", "xyz")]
tib[, c("abc", "xyz")]
```
Again, the tibble outputs a tibble.

**4. Practice referring to non-syntactic names in the following data frame by:**  

  1. Extracting the variable called 1.   
  2. Plotting a scatterplot of 1 vs 2.    
  3. Creating a new column called 3 which is 2 divided by 1.    
  4. Renaming the columns to one, two and three.   
  
```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```

```{r}
# extract variable 1
annoying$`1`
annoying %>% select(`1`)
```

```{r}
# plot scatterplot of 1 vs 2
annoying %>% 
  ggplot(aes(`1`, `2`)) + 
  geom_point()
```

```{r}
# create column called 3, which is 2 divided by 1 
(annoying <- annoying %>% 
  mutate(
    `3` = `2`/`1`
  ))
```

```{r}
# rename columns to one, two, and three
annoying %>% 
  rename(
    one = `1`, 
    two = `2`, 
    three = `3`
  )
```

**5. What does `tibble::enframe()` do? When might you use it?**   

`enframe()` creates tibbles from vectors. This would be useful if you had multiple vectors you wanted to combine together into one data frame/tibble. 

## Chapter 11 

### Exercises 

#### 11.2.2 

**1. What function would you use to read a file where fields were separated with “|”?**   
```{r}
read_delim("a|b|c
1|2|3
4|5|6", delim = "|")
```

**4. Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like `"` or `'`. By default, `read_csv()` assumes that the quoting character will be `"`. What argument to `read_csv()` do you need to specify to read the following text into a data frame?**   
```{r}
read_csv("x,y\n1,'a,b'", quote = "'")
```

**5. Identify what is wrong with each of the following inline CSV files. What happens when you run the code?** 
```{r}
# missing a third column header 
read_csv("a,b\n1,2,3\n4,5,6")

# has three column headers, but only two values in the first row and four values in the second row 
read_csv("a,b,c\n1,2\n1,2,3,4")

# missing a value for the second column 
read_csv("a,b\n\"1")

# not sure what they were going for here? maybe need quotes around values to have "a,b" as one value?
read_csv("a,b\n1,2\na,b")

# use read_csv2?
read_csv("a;b\n1;3")
```

#### 11.3.5

**1. What are the most important arguments to `locale()`?**    
Most important seems subjective, but there are definitely a lot of useful settings in here!

**2. What happens if you try and set `decimal_mark` and `grouping_mark` to the same character? What happens to the default value of `grouping_mark` when you set `decimal_mark` to `“,”`? What happens to the default value of `decimal_mark` when you set the `grouping_mark` to `“.”`?**    
```{r}
# locale(decimal_mark = ".", grouping_mark = ".")
# "decimal_mark and grouping_mark must be different)

parse_number("$123.456,50", locale = locale(decimal_mark = ","))
parse_number("$123.456,50", locale = locale(grouping_mark = "."))
```

`decimal_mark` and `grouping_mark` will just automatically switch with each other. 

**3. I didn’t discuss the `date_format` and `time_format` options to `locale()`. What do they do? Construct an example that shows when they might be useful.**    

`date_format` sets the default date format and `time_format` sets the default time format.  
```{r}
parse_date("05 January 2010", locale = locale(date_format = "%d %B %Y"))
parse_time("1330", locale = locale(time_format = "%H%M"))
```

**7. Generate the correct format string to parse each of the following dates and times:** 
```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"
```

```{r}
parse_date(d1, "%B %d, %Y")
parse_date(d2, "%Y-%b-%d")
parse_date(d3, "%d-%b-%Y")
parse_date(d4, "%B %d (%Y)")
parse_date(d5, "%m/%d/%y")
parse_time(t1, "%H%M")
parse_time(t2, "%I:%M:%OS %p")
```

